---
title: "ODD Car Sales Assignment"
author:
      - Anılcan Atik
      - Dost Karaahmetli
      - Kutay Akalın
      - Tunahan Kılıç
      
date: "November 5th, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ODD Dataset

This is a worksheet of ODD Assignment.The dataset for this project includes car sales numbers from the beginning of 2016 through September 2019.

## Merging Datasets
We begin the project by reading the .rds files that we create through the car sales Excels that we download from the ODD website. For sales_2019, data gathering and celaning code given: (for all years, same code structure was used)


------

library(tidyverse)
library(readxl)

#2018 Sales Data

jan_2018 <- read_excel("2018.01.xlsx", skip=7,col_names=FALSE)
feb_2018 <- read_excel("2018.02.xlsx", skip=7,col_names=FALSE)
mar_2018 <- read_excel("2018.03.xlsx", skip=7,col_names=FALSE)
apr_2018 <- read_excel("2018.04.xlsx", skip=7,col_names=FALSE)
may_2018 <- read_excel("2018.05.xlsx", skip=7,col_names=FALSE)
jun_2018 <- read_excel("2018.06.xlsx", skip=7,col_names=FALSE)
jul_2018 <- read_excel("2018.07.xlsx", skip=7,col_names=FALSE)
aug_2018 <- read_excel("2018.08.xlsx", skip=7,col_names=FALSE)
sep_2018 <- read_excel("2018.09.xlsx", skip=7,col_names=FALSE)
oct_2018 <- read_excel("2018.10.xlsx", skip=7,col_names=FALSE)
nov_2018 <- read_excel("2018.11.xlsx", skip=7,col_names=FALSE)
dec_2018 <- read_excel("2018.12.xlsx", skip=7,col_names=FALSE)


#Slicing unnecessary rows from data

jan_2018 <- jan_2018 %>% slice(-c(44,45))
feb_2018 <- feb_2018 %>% slice(-c(44,45))
mar_2018 <- mar_2018 %>% slice(-c(44,45))
apr_2018 <- apr_2018 %>% slice(-c(43,44))
may_2018 <- may_2018 %>% slice(-c(43,44))
jun_2018 <- jun_2018 %>% slice(-c(43,44))
jul_2018 <- jul_2018 %>% slice(-c(43,44))
aug_2018 <- aug_2018 %>% slice(-c(43,44))
sep_2018 <- sep_2018 %>% slice(-c(43,44))
oct_2018 <- oct_2018 %>% slice(-c(43,44))
nov_2018 <- nov_2018 %>% slice(-c(43,44))
dec_2018 <- dec_2018 %>% slice(-c(43,44))


#Changing col names

dfs <- c("jan_2018", "feb_2018", "mar_2018", "apr_2018", "may_2018", "jun_2018", "jul_2018", "aug_2018", "sep_2018", "oct_2018", "nov_2018", "dec_2018")


for(df in dfs) {
  df.tmp <- get(df)
  names(df.tmp) <- c("brand_name","auto_dom","auto_imp","auto_total","comm_dom","comm_imp","comm_total","total_dom","total_imp","total_total") 
  assign(df, df.tmp)
}


#Creating Month and Year Column

jan_2018 <- jan_2018 %>% mutate_if(is.numeric,funs(ifelse(is.na(.),0,.))) %>% mutate(year = 2018, month = 1)
feb_2018 <- feb_2018 %>% mutate_if(is.numeric,funs(ifelse(is.na(.),0,.))) %>% mutate(year = 2018, month = 2)
mar_2018 <- mar_2018 %>% mutate_if(is.numeric,funs(ifelse(is.na(.),0,.))) %>% mutate(year = 2018, month = 3)
apr_2018 <- apr_2018 %>% mutate_if(is.numeric,funs(ifelse(is.na(.),0,.))) %>% mutate(year = 2018, month = 4)
may_2018 <- may_2018 %>% mutate_if(is.numeric,funs(ifelse(is.na(.),0,.))) %>% mutate(year = 2018, month = 5)
jun_2018 <- jun_2018 %>% mutate_if(is.numeric,funs(ifelse(is.na(.),0,.))) %>% mutate(year = 2018, month = 6)
jul_2018 <- jul_2018 %>% mutate_if(is.numeric,funs(ifelse(is.na(.),0,.))) %>% mutate(year = 2018, month = 7)
aug_2018 <- aug_2018 %>% mutate_if(is.numeric,funs(ifelse(is.na(.),0,.))) %>% mutate(year = 2018, month = 8)
sep_2018 <- sep_2018 %>% mutate_if(is.numeric,funs(ifelse(is.na(.),0,.))) %>% mutate(year = 2018, month = 9)
oct_2018 <- oct_2018 %>% mutate_if(is.numeric,funs(ifelse(is.na(.),0,.))) %>% mutate(year = 2018, month = 10)
nov_2018 <- nov_2018 %>% mutate_if(is.numeric,funs(ifelse(is.na(.),0,.))) %>% mutate(year = 2018, month = 11)
dec_2018 <- dec_2018 %>% mutate_if(is.numeric,funs(ifelse(is.na(.),0,.))) %>% mutate(year = 2018, month = 12)



#Combined 2018 Sales Data

sales_2018 <- rbind(jan_2018, feb_2018, mar_2018, apr_2018, may_2018, jun_2018, jul_2018, aug_2018, sep_2018, oct_2018, nov_2018, dec_2018)


saveRDS(sales_2018,file="sales_2018.rds")


------


```{r warning=FALSE,message=FALSE}
library(tidyverse)
library(ggplot2)
library(zoo)
library(scales)
library(formattable)
library(knitr)
```


```{r readRDS}
sales_2016 <- readRDS(url("https://github.com/pjournal/mef03g-spo-R-ify/blob/master/sales_2016.rds?raw=true"))
sales_2017 <- readRDS(url("https://github.com/pjournal/mef03g-spo-R-ify/blob/master/sales_2017.rds?raw=true"))
sales_2018 <- readRDS(url("https://github.com/pjournal/mef03g-spo-R-ify/blob/master/sales_2018.rds?raw=true"))
sales_2019 <- readRDS(url("https://github.com/pjournal/mef03g-spo-R-ify/blob/master/sales_2019.rds?raw=true"))
```

```{r combined_sales}
combined_sales <- rbind(sales_2016, sales_2017, sales_2018, sales_2019)
combined_sales <- combined_sales %>% filter(!str_detect(brand_name,"ODD"))
combined_sales$brand_name <- str_replace_all(combined_sales$brand_name,"ASTON MARTÄ°N","ASTON MARTIN")
```

We should create the date column for analysis
```{r date column}
combined_sales$YearMonth <- as.Date(as.yearmon(paste(combined_sales$year, " ", combined_sales$month), "%Y %m"))
print(combined_sales, width = Inf)
```


### Monthly car sales from 2016 to September 2019 by brand names
```{r monthly car sales by brand}
sales_byperiod <- combined_sales %>% group_by(YearMonth, brand_name) %>% summarise(total_sales = sum(total_total)) %>% arrange(desc(total_sales)) %>% top_n(1)
sales_byperiod
ggplot(sales_byperiod, aes(x = YearMonth, y = total_sales, fill = brand_name)) + geom_bar(stat = "identity")
```

### Total monthly sales in Turkey
```{r message = FALSE}
monthly_sales <- combined_sales %>% group_by(YearMonth) %>% summarise(total_sales = sum(total_total)) %>% arrange(desc(total_sales))
monthly_sales
ggplot(monthly_sales, aes(x = YearMonth, y = total_sales, color = total_sales)) + geom_point() + geom_smooth()
```

### Total number of car sales by brand from 2016 to Sep.2019
```{r message = FALSE, fig.height=10, fig.width=15, fig.asp= .60}
total_car_numbers <- combined_sales %>% group_by(brand_name) %>% summarise(total_total = sum(total_total)) %>% arrange(desc(total_total))
total_car_numbers
ggplot(total_car_numbers, aes(x = brand_name, y = total_total, size = total_total, color = brand_name)) + geom_point() + theme(axis.text.x = element_text(angle = 90))
```

### Ratio of the number of cars in domestic status to the total number of cars per month
```{r message = FALSE}
domestic_percentage <- combined_sales %>% group_by(YearMonth) %>% summarise(total_dom = sum(total_dom), total_total = sum(total_total)) %>% mutate(percent_total_dom = total_dom/total_total) %>% arrange(desc(percent_total_dom))
domestic_percentage
ggplot(domestic_percentage, aes(x = YearMonth, y = percent_total_dom, size = percent_total_dom, color = percent_total_dom)) + geom_point() + scale_color_gradientn(colours = rainbow(2))
```

### Top 5 Market Share Percentage 2016-2019
```{r message = FALSE}
top5_market_share <- combined_sales %>% group_by(year, brand_name) %>% summarise(total_sales = sum(total_total)) %>% mutate(percent = total_sales / sum(total_sales)) %>% top_n(5, percent)
top5_market_share
ggplot(top5_market_share, aes(x = year, y = percent, color = brand_name)) + geom_point() + geom_line()
```

### Top 20 Firms Holding the Market Share between 2016-2019
```{r message = FALSE}
combined_sales$time<-as.Date(as.yearmon(paste(combined_sales$year, " ", combined_sales$month), "%Y %m"))
M_S<-combined_sales%>%select(brand_name,year,total_total)%>%group_by(brand_name)%>%mutate(t_total=sum(total_total))
M_S1<-M_S[!duplicated(M_S$brand_name), ]
sum_total<-sum(M_S1$t_total)
M_S2<-M_S1%>%arrange(desc(t_total))%>%head(20)
ggplot(M_S2,aes(x=reorder(brand_name,-t_total),y=t_total/sum_total,color=brand_name))+geom_col(aes(fill=brand_name))+theme(axis.text.x = element_text(angle = 90, hjust = 1))+labs(title = "2016-2019 Total Market Share in Percentage",y="Market Share in Percentage",x="Company Names")
```

### Total Imported Car Shares of Brands between 2016-2019
```{r message = FALSE}
T_I<-combined_sales%>%select(brand_name,total_imp,time)%>%group_by(time)%>% mutate(perc_total_imp = total_imp/sum(total_imp))%>%top_n(10)
ggplot(T_I,aes(x=time,y=total_imp,color=brand_name))+geom_col(aes(fill=brand_name))+scale_x_date(labels = date_format("%Y %m"))+labs(title = "2016-2019 Total Imports",y="Total Imports",x="Time")
```

### Total Domestic Car Shares of Brands between 2016-2019
```{r message = FALSE}
T_D<-combined_sales%>%select(brand_name,total_dom,time)%>%group_by(time)%>% mutate(perc_total_dom = total_dom/sum(total_dom))%>%top_n(10)  
ggplot(k,aes(x=time,y=total_dom))+geom_col(aes(fill=brand_name))+scale_x_date(labels = date_format("%Y %m"))+labs(title = "2016-2019 Total Domestic",y="Total Domestic",x="Time")
```

### Total Domestic, Imported Car Sales in Months Beween 2016-2019
```{r message = FALSE}

slct_1<-combined_sales %>% group_by(year,month) %>% summarise(sum_dom=sum(total_dom),sum_imp=sum(total_imp),sum_total=sum(total_total),sum_dom) %>% ungroup()

formattable(slct_1)
```

### Bar Graph of 4 Year Total Car Sales in Months
```{r message = FALSE}

levels_month= c("1","2","3","4","5","6","7","8","9","10","11","12")

ggplot(slct_1,aes(x=ordered(as.character(month),levels=levels_month),y=sum_total,fill=ordered(year,levels=c("2016","2017","2018","2019")))) +geom_bar(stat="identity",position="dodge") + labs(y="Total Car Sales",x="Months",fill="")

```
