---
title: "Spotify Project: A Summary of the Data and Explanations"
author: 
    - _Anılcan_ _Atik_
    - _Dost_ _Karaahmetli_
    - _Kutay_ _Akalın_
    - _Tunahan_ _Kılıç_
      
date: "November 17th, 2019"
output: 
  html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Explanation

Our data obtained directly from Spotify Web API. For API connection, we created “Client ID” and “Client Secret” from [Spotify for Developers Website](https://developer.spotify.com/documentation/web-api/). For this purpose, **"spotifyr"** package used for making connection.

### Accessing Spotifty Web API
```{r message=FALSE, warning=FALSE, include=FALSE}
library(httpuv)
library(spotifyr)
library(tidyverse)
library(knitr)
library(lubridate)
library(ggplot2)
Sys.setenv(SPOTIFY_CLIENT_ID = 'f5adea41ba0c4184a3d15e9960b4a0c2')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '3b5362e9b6a44ea6a774911ae7700334')
access_token <- get_spotify_access_token()
```

When connection is made successfully, we can access lots of difrent type data such as aritst, albums, tracks, user profile etc. Here is the Spotify API [References](https://developer.spotify.com/documentation/web-api/reference/). In our project, we will usually use playlist, artist and track data.

#### Track

[Documentation on Tracks Data](https://developer.spotify.com/documentation/web-api/reference/tracks/)

#### Playlist 

[Documentation on Playlists Data](https://developer.spotify.com/documentation/web-api/reference/playlists/)

#### Artist

[Documentation on Artists Data](https://developer.spotify.com/documentation/web-api/reference/artists/)


## Initial Analysis
### Turkey Top 50 Analysis

```{r}
turkey_top_50_id="37i9dQZEVXbIVYVBNw9D5K"
turkey_top50_audio_features <- get_playlist_audio_features("spotifycharts",turkey_top_50_id)
glimpse(turkey_top50_audio_features)
```



### Adding Sentiments in Each Track

The purpose of this function named _"classify_track_sentiment"_ is important for us to work primarily to reveal the mood of songs and song lists along these lines. Energy and valence are two important factors in terms of interpreting emotion in music. The variations of these two factors, which have values between 0 and 1, in this range determine the songs to be turbulent/angry, happy/joyful, sad/depressing and chill/peaceful.

According to [Get Audio Features for a Track](https://developer.spotify.com/documentation/web-api/reference/tracks/get-audio-features/), explanations of the corresponding factors are as follows.

#### Energy
Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity. Typically, energetic tracks feel fast, loud, and noisy. For example, death metal has high energy, while a Bach prelude scores low on the scale. Perceptual features contributing to this attribute include dynamic range, perceived loudness, timbre, onset rate, and general entropy. 

#### Valence
A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track. Tracks with high valence sound more positive (e.g. happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g. sad, depressed, angry).
```{r}
classify_track_sentiment <- function(valence, energy) {
    if (is.na(valence) | is.na(energy)) {
        return(NA)
    }
    else if (valence >= .5) {
        if (energy >= .5) {
            return('Happy/Joyful')
        } else {
            return('Chill/Peaceful')
        }
    } else {
        if (energy >= .5) {
            return('Turbulent/Angry')
        } else {
            return('Sad/Depressing')
        }
    }
}
track_sentiment = c()
for (i in 1:50){
  track_sentiment[i]= classify_track_sentiment(turkey_top50_audio_features[[15]][[i]],turkey_top50_audio_features[[7]][[i]])
  
}
turkey_top50_audio_features<-cbind(turkey_top50_audio_features,track_sentiment)
```

### Selection of Necessary Columns for Analysing

```{r}
track_id="6bBnnrknLbDoOCUdKMkmnq"
track_audio_analyse_selection<- turkey_top50_audio_features %>% select(track.name,track.id,track.artists,track.album.images,track.album.release_date,track.popularity,danceability:tempo,track_sentiment,track.duration_ms)
head(track_audio_analyse_selection,width=10)
  glimpse(track_audio_analyse_selection)
```

### Artists in Top 50
```{r}
artists=c()
for (i in 1:50){
  
  artists <- c(artists, track_audio_analyse_selection[[3]][[i]][[3]])
  
}
#Changing artists vector as a dataframe and sorting bypearing freaquency in top 50 music
artist_df<-as.data.frame(table(artists)) %>% arrange(desc(Freq))
head(artist_df, 15L)
```


### Sentiment Counts in Playlist
```{r}
sent_count <- track_audio_analyse_selection %>% group_by(track_sentiment) %>% count()
sent_count
```

```{r}
ggplot(sent_count, aes(x=track_sentiment, y=n, fill=track_sentiment)) +
  geom_bar(stat="identity")
```


```{r}
ggplot(track_audio_analyse_selection,aes(x = valence, y = energy, color = track_sentiment)) + geom_point() +
  labs(color = "", title = "Sentiment Analysis of Turkey Top 50 Chart") +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 1)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
  geom_label(aes(x = 0.25, y = 0.97, label = "Turbulent/Angry"), label.padding = unit(2, "mm"),  fill = "darkgrey", color="white") +
  geom_label(aes(x = 0.75, y = 0.97, label = "Happy/Joyful"), label.padding = unit(2, "mm"), fill = "darkgrey", color="white") +
  geom_label(aes(x = 0.25, y = 0.03, label = "Sad/Depressing"), label.padding = unit(2, "mm"),  fill = "darkgrey", color="white") +
  geom_label(aes(x = 0.75, y = 0.03, label = "Chill/Peaceful"), label.padding = unit(2, "mm"), fill = "darkgrey", color="white") +
  geom_segment(aes(x = 1, y = 0, xend = 1, yend = 1)) +
  geom_segment(aes(x = 0, y = 0, xend = 0, yend = 1)) +
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 0)) +
  geom_segment(aes(x = 0, y = 0.5, xend = 1, yend = 0.5)) +
  geom_segment(aes(x = 0.5, y = 0, xend = 0.5, yend = 1)) +
  geom_segment(aes(x = 0, y = 1, xend = 1, yend = 1)) 
  
```