---
title: "Spotify Project"
author:
      - _Anılcan_ _Atik_
      - _Dost_ _Karaahmetli_
      - _Kutay_ _Akalın_
      - _Tunahan_ _Kılıç_
      
date: "November 17th, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Explanation

Our data obtained directly from Spotify Web API. For API connection, we created “Client ID” and “Client Secret” from Spotify for Developers [Website]https://developer.spotify.com/documentation/web-api/). For this purpose, "spotifyr" package used for making connection.

### Accessing Spotifty Web API
```{r message=FALSE}
library(httpuv)
library(spotifyr)
library(tidyverse)
library(knitr)
library(lubridate)


Sys.setenv(SPOTIFY_CLIENT_ID = 'f5adea41ba0c4184a3d15e9960b4a0c2')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '3b5362e9b6a44ea6a774911ae7700334')

access_token <- get_spotify_access_token()
```

When connection is made successfully, we can access lots of difrent type data such as aritst, albums, tracks, user profile etc. Here is the Spotify API [References](https://developer.spotify.com/documentation/web-api/reference/). In our project, we will usually use playlist, artist and track data.

#### Track

[Şarkı verilerinin açılanması](https://developer.spotify.com/documentation/web-api/reference/tracks/)

#### Playlist 

[Plasylist verilerinin açılanması](https://developer.spotify.com/documentation/web-api/reference/playlists/)

#### Artist

[Sanatçı verilerinin açılanması](https://developer.spotify.com/documentation/web-api/reference/artists/)


## Initial Analysis
### Turkey Top 50 Analysis

```{r}

turkey_top_50_id="37i9dQZEVXbIVYVBNw9D5K"

turkey_top50_audio_features <- get_playlist_audio_features("spotifycharts",turkey_top_50_id)
turkey_top50_audio_features
```



#### Adding Sentiments in Each Track
```{r}

classify_track_sentiment <- function(valence, energy) {
    if (is.na(valence) | is.na(energy)) {
        return(NA)
    }
    else if (valence >= .5) {
        if (energy >= .5) {
            return('Happy/Joyful')
        } else {
            return('Chill/Peaceful')
        }
    } else {
        if (energy >= .5) {
            return('Turbulent/Angry')
        } else {
            return('Sad/Depressing')
        }
    }
}

track_sentiment = c()
for (i in 1:50){
  track_sentiment[i]= classify_track_sentiment(turkey_top50_audio_features[[15]][[i]],turkey_top50_audio_features[[7]][[i]])
  
}

turkey_top50_audio_features<-cbind(turkey_top50_audio_features,track_sentiment)
```

#### Selection of Necessary Columns for Analysing

```{r}

track_id="6bBnnrknLbDoOCUdKMkmnq"
track_audio_analyse_selection<- turkey_top50_audio_features %>% select(track.name,track.id,track.artists,track.album.images,track.album.release_date,track.popularity,danceability:tempo,track_sentiment,track.duration_ms)

print(track_audio_analyse_selection,width=10)
  glimpse(track_audio_analyse_selection)


```

#### Artists in Top 50
```{r}

artists=c()
for (i in 1:50){
  
  artists <- c(artists, track_audio_analyse_selection[[3]][[i]][[3]])
  
}

#Changing artists vector as a dataframe and sorting bypearing freaquency in top 50 music
artist_df<-as.data.frame(table(artists)) %>% arrange(desc(Freq))
artist_df
```


#### Sentiment Counts in Playlist
```{r}

track_audio_analyse_selection %>% group_by(track_sentiment) %>% count()
```




