---
title: "BES Assignment"
author:
      - _Anılcan_ _Atik_
      - _Dost_ _Karaahmetli_
      - _Kutay_ _Akalın_
      - _Tunahan_ _Kılıç_
      
date: "November 8th, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.BES Dataset

This is a worksheet of BES Assignment.The dataset for this project includes key indicators of Individual Pension System (BES) from the January 2017 through July 2019.

## 2.Preparing and Cleaning Dataset 
We begin the project by reading the excel files that we download from the [EGM website](https://www.egm.org.tr/bilgi-merkezi/istatistikler/). After the "Rapor Tarihi" filter was removed, raw data was created in the excel sheet. This sheet was uploaded to [spo'R'ify Github Page](https://github.com/pjournal/mef03g-spo-R-ify/tree/master/BES%20Assignment) for further analysis. Downloading file and cleaning process has given below:


```{r warning=FALSE,message=FALSE}
library(tidyverse)
library(readxl)
library(bsts)
#Downloading the excelel file
tmp<-tempfile(fileext=".xlsx")
download.file("https://github.com/pjournal/mef03g-spo-R-ify/blob/master/BES%20Assignment/BES_2017_07.2019.xlsx?raw=true",mode="wb",destfile=tmp)
                      
bes_data <-readxl::read_excel(tmp,skip=2,col_names=FALSE)
file.remove(tmp)
#cleaining process
names(bes_data) <- c("sirket_adi", "katilimci_sayisi", "katilimci_fon_TL", "devlet_katkisi_TL", "katki_payi_TL", "tarih", "ss_emekli_sayisi", "ss_bireysel_sozlesme", "ss_grup_bireysel_sozlesme", "ss_isveren_grup_sertifika", "ss_toplam", "yy_bireysel_TL", "yy_grup_bireysel_TL", "yy_isveren_grup_TL", "yy_toplam_TL")
bes_data$tarih <-  format(as.Date(bes_data$tarih), "%Y-%m-%d")
bes_data$tarih <- as.Date(bes_data$tarih)
#To use monthly data, we need to get the last day of each month. With this code, we select the last day of the month.
#In pipe, "filter(tarih %in% lastday)" can be used.
lastday <- unique(LastDayInMonth(bes_data$tarih))
bes_data <- bes_data %>% mutate_if(is.numeric,funs(ifelse(is.na(.),0,.)))
options(scipen=999)
```



## 3.Loding Libraries and Civilized Data Display
```{r warning=FALSE,message=FALSE}
library(tidyverse)
library(ggplot2)
library(zoo)
library(scales)
library(formattable)
library(knitr)
library(bsts)
print(bes_data,width=Inf)
```

## 4.Exploratory Data Analysis

### 4.1.Total Fund Values for Each Firm Yearly
```{r fig.height=10, fig.width=10, fig.asp= .60}
bes_data_monthly<- bes_data %>% filter(tarih %in% lastday) %>% mutate (year = format(tarih,"%Y"),month=format(tarih,"%B")) 
slct_1<-bes_data_monthly %>% filter (month == "Temmuz")
slct_1
ggplot(slct_1,aes(x=sirket_adi,y=katilimci_fon_TL ,fill=ordered(year))) +geom_bar(stat="identity",position="dodge") + labs(y="Total Funds",x="Firms",fill="") + theme(axis.text.x = element_text(angle = 90, hjust = 1))+scale_y_continuous(labels=comma)
```


### 4.2.Ratio of Number of Retired Participants to Total Number of Participants by Each Firm
```{r fig.height=10, fig.width=10, fig.asp= .60, message=FALSE}
bes_katilimci <- bes_data %>% filter(tarih %in% lastday) %>% group_by(sirket_adi, tarih) %>% summarise(emekli_toplam = sum(ss_emekli_sayisi), sozlesme_toplam = sum(ss_toplam), emekli_oran = emekli_toplam / sozlesme_toplam)

bes_katilimci

ggplot(data = bes_katilimci, mapping = aes(x = tarih, y = emekli_oran, color = sirket_adi, group = 2)) +
  geom_line() +
  facet_wrap(facets =  vars(sirket_adi)) +
  labs(title = "Ratio of Retired Participants to Total Participants", x = "Date", y = "Ratio of Retired Participants ", color = " ") +
  theme(axis.text = element_text( size = 8 ),
        axis.text.x = element_text( size = 8 ), 
        axis.title = element_text( size = 10, face = "bold" ),
         legend.position="none",
         # The new stuff
         strip.text = element_text(size = 8))
```


### 4.3.Top 5 Companies
```{r Top 5 Companies}
# Table: Top 5 Companies by Participants:
top_5_by_katilimci_sayisi <- bes_data %>% 
  group_by(sirket_adi) %>%
  summarise(total_ks=sum(katilimci_sayisi)) %>%
  arrange(desc(total_ks)) %>%
  top_n(5, total_ks) 
print(top_5_by_katilimci_sayisi)


# Table: Top 5 Companies by Number of Contracts:
top_5_by_sozlesme_sayisi <- bes_data %>% 
  group_by(sirket_adi) %>%
  summarise(total_ss=sum(ss_toplam)) %>%
  arrange(desc(total_ss)) %>%
  top_n(5, total_ss) 
print(top_5_by_sozlesme_sayisi)


# Table: Top 5 Companies by Market Share (Mean Participant Funds):
top_5_by_market_share <- bes_data %>% 
  group_by(sirket_adi) %>%
  summarise(mean_of_fundsize=mean(katilimci_fon_TL, na.rm = TRUE)) %>%
  mutate(mshare = percent(mean_of_fundsize / sum(mean_of_fundsize))) %>%
  top_n(5, mshare) 
print(top_5_by_market_share)


# Plot: Top 5 Companies by Market Share (Mean Participant Funds):
ggplot(data = top_5_by_market_share, aes(x=sirket_adi, y=mshare, fill=sirket_adi)) +
  geom_bar(stat="identity")+
  coord_polar()+
  theme(legend.position = "right", axis.text.x = element_text(angle = 0))+
  geom_text(aes(y = mshare , label = mshare))+
  labs(title = "Market shares of top 5 companies", x="", y="")
```
